---
title: "external combination validation panel lib_generation"
author: "Niklas Rindtorff"
output: 
   BiocStyle::html_document2:
    code_folding: hide
    df_print: paged
    toc_depth: 3
    toc_float: true
    fig_height: 6
    fig_width: 9.5
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load data
##Load libraries  

```{r, warning=FALSE, message=FALSE}
library(readxl)
library(tidyverse)
library(readr)
```

#Load Examples
As before the library dispensing is divided into 3 parts. 
As templates I load example files from the last library preparation round. 


```{r, warning=FALSE}
robot_cmd_1_dilu <- read_csv("~/promise/local_data/layouts/robotic_files/l08_ccp_v1/Solve_Compounds/Work_list_Step_3.csv")
robot_cmd_2_drug <- read_csv("~/promise/local_data/layouts/robotic_files/l08_ccp_v1/robot_clinicalcancerpanel_worklist.csv") %>%
  rename(drug = `Drug Name`)
robot_cmd_3_ctrl <- read_csv("~/promise/local_data/layouts/robotic_files/l08_ccp_v1/Worklist_DMSO_Stauro_spotting.csv")
```

Example of step 1

```{r}
robot_cmd_1_dilu %>% head()
```

This example illustrates how compounds are sampled in the 2. step. 

```{r}
robot_cmd_2_drug %>% 
  filter(drug == "Bortezomib")
```

Example of step 3

```{r}
robot_cmd_3_ctrl %>% head()
```

#Prepare protocol 
The current protocol differs conserideably from former approaches as way fewer compounds will be dispensed but with a higher complexity. 

##Step 1
In this setup step 1 will be conducted manual or with already available tools from Beckman. 
A deepwell plate will be filled and a rowwise dilution will be applied on some columns. 

##Step 2
This protocol will take 2*(384-60ctrls) commands to dispense the combinaitions.

First I build a schematic tibble.
```{r}
step2 <- tibble(conc1 = paste(c(1:6), c("1-max", c(2:4), "5-min", "6-DMSO"), sep ="_")) %>% 
  mutate(conc2 = conc1) %>% 
  cbind(.,drug_pair = c("AB", "CD", "EF")) %>% #hack around vector length restrictions 
  #I name the pairs by the rows in which the compounds will be pipetted
  cbind(., rep = c(1:3)) %>% 
  expand(conc1, conc2, drug_pair, rep) %>%
  #Now I assign source wells etc. 
  separate(drug_pair, c("row1", "row2"), remove = FALSE, sep =1) %>% 
  separate(conc1, c("col1", "conc1"), sep = "_") %>%
  separate(conc2, c("col2", "conc2"), sep = "_") %>%
  arrange(drug_pair,rep) %>% 
  unite(well1, row1, col1, sep = "") %>% 
  unite(well2, row2, col2, sep = "")

head(step2)
```

Now I reformat the table to end up with a robot command file

First I take a look again at my target column names 
```{r}
colnames(robot_cmd_2_drug)
```

Now let's go. 
First I need to give every combination a destination.well

```{r}
step2 <- step2 %>% 
  mutate(destination.well = c(robot_cmd_2_drug$destination.well[1:324]) %>% sample()) %>% #I sample my destination wells from a plate layout that I used before in a random order. However, this library comprises less compound wells than the reference library that I showed before. Therfore I take the first 324 wells from this vector only.
  #dplyr::select(-conc1, -conc2) %>%
  gather(well, number, -drug_pair, -rep, -destination.well, -conc1, -conc2) %>% 
  mutate(Destination_Plate = "384_Master",
         Source_Position = "Master_Library_DWPlate",
         volume = 30) %>% 
  rename(source.well = number,
         pipetting_iteration = well) %>% 
  unite(concentration_pair, conc1, conc2)
  
colnames(step2)
  
```

I take a final look at the table. 

```{r}
step2
```


###QC 

I check how often every destination well is present in the dataset

```{r}
test <- step2 %>% group_by(destination.well) %>% 
  summarise(n =n()) %>% filter(n ==2) %>% nrow()

test == 324
```

I estimate the amount of compound which I need per well in the source plate

```{r}
test <- step2 %>% group_by(source.well, volume) %>% 
  summarise(n = n()) %>% mutate(vol = volume * n) %>% .$vol

test <- (test > 580) %>% sum() #600 ul minus pipetting mistakes
test == 0
```

###Save protocol
I save the step2 protocol as .csv

```{r}
write.csv(step2, file = "l12_ecvp_robot_cmd_2_drug_combi.csv")
```

##Step3
Now I add the controls to the final library. 

```{r}
step3 <- tibble(substance = c(rep(c("DMSO"), each = c(40)), rep(c("Staurosporine"), each = c(20)))) %>% 
  cbind(., source.well = c(rep(c("A7", "B7", "C7", "D7"), each = c(10)), rep(c("A8", "B8"), each = c(10)))) %>%
#I add the compound source wells manually as not all susbtance can be drawn form one well.
  mutate(destination.well = c(robot_cmd_2_drug$destination.well[325:nrow(robot_cmd_2_drug)], 
                              robot_cmd_3_ctrl$destination.well) %>% sample()) %>%
  mutate(Destination_Plate = "384_Master",
         Source_Position = "Master_Library_DWPlate",
         volume = 60)
 
step3
```

###QC

no destination well should be part of the step2 protocol

```{r}
test <- step3$destination.well %in% step2$destination.well %>% sum()
test == 0
```

The volume should be as expected.

```{r}
test <- step3 %>% group_by(source.well, volume) %>% 
  summarise(n = n()) %>% mutate(vol = n* volume) %>% .$vol

test <- (test > 720) %>% sum() #600 ul minus pipetting mistakes
test == 0
```

###Save protocol 

I save the step3 protocol as .csv

```{r}
write.csv(step3, file = "l12_ecvp_robot_cmd_3_ctrl.csv")
```


#Final QC

```{r}
nrow(step2) + nrow(step3) == (324*2 + (384-324))
```

```{r, eval = FALSE}

library(platetools)
test <- tibble(well =  )

step2 %>% unite(id, concentration_pair, drug_pair) %>% dplyr::select(id, destination.well) %>

raw_map(data = ,
        well = df$well,
        plate = 384)
```

