---
title: "Tidy and collection follow-up data"
output:
  pdf_document: default
  html_notebook: default
---

```{r}
library(tidyverse)
library(here)

library(cowplot)
library(pheatmap)
library(splots)
library(patchwork)

library(synergyfinder)
library(ggsignif)
library(broom)
library(ggrepel)
library(readxl)
library(platetools)
library(PharmacoGx)
#library(ggridges)
library(reshape2)
```

## load data

```{r}
#source(here::here("src/data/make_ctg_data.R"))
#source(here::here("src/data/"))

ctg_data <- read_csv(here::here("data/processed/ctg_data.csv"))
ctg_loess <- read_csv(here::here("data/processed/ctg_loess.csv"))
ctg_norm <- read_csv(here::here("data/processed/ctg_norm.csv"))
syn_table <- read_csv(here::here("data/processed/synergy_input.csv"))
syn_scores <- read_csv(here::here("data/processed/synergy_scores.csv"))
```


First, I get an overview of the dataset. 

```{r}
pdf(file = "overview_valid_raw.pdf", width = 15, height = 4, compress = FALSE)
ctg_data %>% 
  select(id, pcount) %>% arrange(id) %>% split(., .$id) %>% lapply(., function(f) f$pcount %>% as.vector) %>%
  plotScreen(., 10, do.names = TRUE, do.legend = FALSE)
dev.off()

ctg_data %>% 
  select(id, pcount) %>% arrange(id) %>% split(., .$id) %>% lapply(., function(f) f$pcount %>% as.vector) %>%
  plotScreen(., 10, do.names = TRUE, do.legend = FALSE)
```


Now I focus on column and row-wise effects. Row wise effects dominate clearly.

```{r}
ctg_data %>% ggplot(aes(row_num, pcount)) + geom_point() + 
  geom_smooth() +
  facet_wrap(~id, nrow = 4) + theme_bw() + ggtitle("Spatial effects of rows") 

```

```{r}
  #next figure
  ctg_data %>% ggplot(aes(col_num, pcount)) + geom_point() + 
  geom_smooth() +
  facet_wrap(~id, nrow = 4) + theme_bw() + ggtitle("Spatial effects of columns")
```

I applied a conservative loess fit for both columns and rows. 
I plot the results for rows and columns below.

```{r}
ctg_loess %>% ggplot(aes(row_num, pcount_norm)) + geom_point() + 
  geom_smooth() +
  facet_wrap(~id, nrow = 4) + theme_bw() + ggtitle("Spatial effects of rows") 

```

```{r}
  #next figure
  ctg_loess %>% ggplot(aes(col_num, pcount_norm)) + geom_point() + 
  geom_smooth() +
  facet_wrap(~id, nrow = 4) + theme_bw() + ggtitle("Spatial effects of columns")
```

```{r}
pdf(file = "overview_valid_norm.pdf", width = 15, height = 4, compress = FALSE)
ctg_loess %>% 
  select(id, pcount_norm) %>% arrange(id) %>% split(., .$id) %>% lapply(., function(f) f$pcount_norm %>% as.vector) %>%
  plotScreen(., 10, do.names = TRUE, do.legend = FALSE)
dev.off()

ctg_loess %>% 
  select(id, pcount_norm) %>% arrange(id) %>% split(., .$id) %>% lapply(., function(f) f$pcount_norm %>% as.vector) %>%
  plotScreen(., 10, do.names = TRUE, do.legend = FALSE)

```

# QC

Now I look at the distribution of controls in the first screening replicate.

```{r, warning=FALSE}
coi <- c("DMSO")

ctg_loess %>%
  mutate(date = as.character(date)) %>%
  filter(drug_1 %in% coi & drug_2 %in% coi) %>%
  ggplot(aes(line, pcount_norm, color = date)) +
  #geom_violin(aes(group = drug)) +
  geom_jitter(alpha = 0.7, width = 0.1) +
  #geom_boxplot(width = 30) +
  #facet_wrap(~line) +
  theme_cowplot() + 
  coord_flip()

# ctg_loess %>%
#   mutate(line_date_exp = paste0(line, "_", date)) %>%
#   filter(drug %in% coi) %>%
#   filter(combination == TRUE) %>%
#   ggplot(aes(id, pcount_norm, color = drug)) +
#   #geom_violin(aes(group = drug)) +
#   geom_boxplot(alpha = 0.7) + 
#   #facet_grid(line_date_exp ~ Library) + 
#   #facet_wrap(~line) +
#   theme_classic()
```

There are both line and replicate level differences in viability. I normalize on a plate level. Before I move on, I plot the distribution of photnon counts for each plate. 

```{r}
ctg_loess %>%
  ggplot(aes(pcount_norm, fill = date)) +
  #geom_violin(aes(group = drug)) +
  geom_histogram(position="identity", alpha = 0.4) +
  #geom_boxplot(width = 30) +
  facet_wrap(~line) +
  theme_cowplot() + 
  geom_vline(xintercept = 0)
```

I expect the highest concentration of taxanes to be completely lethal. 

```{r}
coi <- c("Docetaxel_DMSO", "DMSO_DMSO")

ctg_loess %>%
  filter(drug %in% coi) %>% 
  filter(conc_1 == 1) %>%
  ggplot(aes(drug_pair_conc, pcount_norm)) + 
  geom_point() + 
  facet_wrap(~ line) + 
  theme_cowplot()
```


I calculate z-factors, as a reference point, I select the highest concentration pair 

```{r, eval = TRUE}
coi <- c("Docetaxel_DMSO", "DMSO_DMSO")

z_df <- ctg_loess %>%
  filter(drug %in% coi) %>%
  filter(!(drug == coi[1] & conc_1 %in% c(3:5))) %>%
  group_by(id, drug, line, date) %>%
  summarise(sd = sd(pcount_norm, na.rm = TRUE),
            mean = mean(pcount_norm, na.rm = TRUE)) %>%
  group_by(id, line, date) %>%
  summarise(zfactor = 1-((3*sum(sd))/abs(range(mean)[1]-range(mean)[2]))) %>%
  mutate(qc = if_else(zfactor < 0.25, FALSE, TRUE))

z_df %>% 
  arrange(zfactor) %>%
  ungroup() %>%
  mutate(id = factor(id, levels = id)) %>%
  ggplot(aes(zfactor, id)) + 
  geom_point(aes(color =  line)) + 
  theme_cowplot() 

# z_df_anno <- z_df %>% ungroup%>% mutate(id = paste0(id, "_", combination)) %>% 
#   dplyr::select(-experiment, -zfactor) %>%
#   as.data.frame() %>% 
#   remove_rownames() %>%
#   column_to_rownames("id")
# 
# z_df %>% ungroup%>% mutate(id = paste0(id, "_", combination)) %>% dplyr::select(id, zfactor) %>% 
#   arrange(zfactor) %>%
#   as.data.frame() %>% 
#   remove_rownames() %>%
#   column_to_rownames("id") %>% 
#   pheatmap(., cluster_cols = FALSE, annotation_row =  z_df_anno)
```

Now I create a correlation plot for both replicates across all plates except the column 13 combination wells. 

```{r}
r <- ctg_loess %>%
  mutate(ctrl = ifelse(drug == "DMSO", "negative", "none")) %>% 
  mutate(ctrl = ifelse(drug == "Docetaxel_DMSO" & conc_1 %in% c(1:2), "positive", ctrl)) %>% 
  #filter(drug %in% coi) %>%
  unite(id, c("line", "well")) %>%
  dplyr::select(id, pcount_norm, replicate) %>% 
  spread(replicate, pcount_norm) %>% 
  rename(rep1 = `1`) %>%
  rename(rep2 = `2`) %>%
  dplyr::select(rep1, rep2) %>% 
  as.matrix() %>%
  cor() %>%
  .["rep1", "rep2"]

ctg_loess %>%
  
  mutate(ctrl = ifelse(drug == "DMSO", "negative", "none")) %>% 
  mutate(ctrl = ifelse(drug == "Docetaxel_DMSO" & conc_1 %in% c(1:2), "positive", ctrl)) %>% 
  #filter(drug %in% coi) %>%
  unite(id, c("line", "well")) %>%
  dplyr::select(id, pcount_norm, replicate) %>% 
  spread(replicate, pcount_norm) %>% 
  rename(rep1 = `1`) %>%
  rename(rep2 = `2`) %>%
  ggplot(aes(rep1, rep2)) + 
  geom_point(alpha = 0.2) +
  #geom_point(alpha = 0.2, aes(color = ctrl)) + 
  #scale_color_manual(values=c("#E69F00", "#999999", "#56B4E9")) +
  geom_abline(intercept = 0, slope = 1) +
  #geom_hline(yintercept = 1) +
  theme_classic() + 
  ggtitle(paste0("r (pearson) = ", r %>% round(2))) + 
  xlab("replicate 1") + 
  ylab("replicate 2") + 
  scale_y_continuous(breaks = c(0, 250000, 500000, 750000)) + 
  coord_fixed() + 
  ggsave("pearson_raw_valid.pdf", width = 4, height = 4)
  
  
```


#Normalization 

At first glance, the majority of the screen is well behaved. I performed plate-wise noramlization to see wether I can also recover plates with differring plates.

I plot the correlation again. 

```{r}
r <- ctg_norm %>%
  mutate(ctrl = ifelse(drug == "DMSO_DMSO", "negative", "none")) %>% 
  mutate(ctrl = ifelse(drug == "Docetaxel_DMSO" & conc_1 %in% c(1:2), "positive", ctrl)) %>% 
  #filter(drug %in% coi) %>%
  unite(id, c("line", "well")) %>%
  dplyr::select(id, viability, replicate) %>% 
  spread(replicate, viability) %>% 
  rename(rep1 = `1`) %>%
  rename(rep2 = `2`) %>%
  dplyr::select(rep1, rep2) %>% 
  as.matrix() %>%
  cor() %>%
  .["rep1", "rep2"]

df <- ctg_norm %>%
  mutate(ctrl = ifelse(drug == "DMSO", "negative", "none")) %>% 
  mutate(ctrl = ifelse(drug == "Docetaxel_DMSO" & conc_1 %in% c(1:2), "positive", ctrl)) %>% 
  #filter(drug %in% coi) %>%
  unite(id, c("line", "well")) %>%
  dplyr::select(id, viability_uncropped, replicate) %>% 
  spread(replicate, viability_uncropped) %>% 
  rename(rep1 = `1`) %>%
  rename(rep2 = `2`) 

df %>% write_csv(here::here("results/plot_data/pearson_norm_valid.csv"))

df %>%
  ggplot(aes(rep1, rep2)) + 
  geom_point(alpha = 0.2) +
  #geom_point(alpha = 0.2, aes(color = ctrl)) + 
  #scale_color_manual(values=c("#E69F00", "#999999", "#56B4E9")) +
  geom_abline(intercept = 0, slope = 1) +
  #geom_hline(yintercept = 1) +
  theme_classic() + 
  ggtitle(paste0("r (pearson) = ", r %>% round(2))) + 
  xlab("replicate 1") + 
  ylab("replicate 2") + 
  scale_y_continuous(breaks = c(0, 0.5, 1, 1.5,2)) + 
  scale_x_continuous(breaks = c(0, 0.5, 1, 1.5,2)) + 
  geom_hline(yintercept = 1) + 
  geom_vline(xintercept = 1) +
  coord_fixed() + 
  ggsave("pearson_norm_valid.pdf", width = 4, height = 4)
  
```

To avoid the simpson paradox, I also plot line-wise correlations.


```{r}
ctg_norm %>%
  dplyr::select(id, viability, well) %>%
  spread(id, viability) %>% 
  dplyr::select(-well) %>% 
  as.matrix() %>% 
  cor() %>% 
  pheatmap::pheatmap()
```

 Synergy calculations are missled in cases where are treatment is actually boosting overall viability. Because of this effect, I set the maximum viability to 1. 
 
```{r}
ctg_norm <- ctg_norm %>% 
  mutate(viability = if_else(viability > 1, 1, viability))
```


# Calculating synergy scores
I consider using the "synergyfinder" package, published by He et al. at the FIMM. 
First I wrangled the data into a standardized input format.

I perform a sanity check after preparing the data for synergy exploration.

```{r}
f <- syn_table %>% filter(block_id == "Everolimus_AZD5363", line == "D013T01", replicate == 1) %>% 
  ungroup() %>% 
  dplyr::select(-replicate)

# trouble shooting
# me = c("ZIP", "HSA", "Bliss", "Loewe")
# m = me[1]
f
```

I visaulize the matrix of responses 

```{r}
syn_table %>% filter(replicate ==1) %>% 
  mutate(conc_r = factor(conc_r),
         conc_c = factor(conc_c)) %>%
  ggplot(aes(conc_r, conc_c, fill = response)) + 
  geom_tile() + 
  scale_fill_viridis_c() + 
  facet_grid(block_id ~ line) + 
  theme_cowplot()+ 
  ggsave("response_overview.pdf", height = 15, width = 17)
```

* Some concentrations of Napa_Suni are missing. I check if some of the data was lost on the way.
* D093N11 is not interpretable and should be removed. 

```{r}
# syn_table shows the missing data
syn_table %>% 
  group_by(block_id) %>% 
  summarise(n = n()) %>% 
  filter(n != max(n))

# ctg_norm is not ok
ctg_data %>% 
  mutate(block_id = paste(drug, sep = "_")) %>% 
  group_by(block_id) %>% 
  summarise(n = n()) %>% 
  count(n)

# I filter n = 130, 104, 416
ctg_data %>% 
  mutate(block_id = paste(drug, sep = "_")) %>% 
  group_by(block_id) %>% 
  summarise(n = n()) %>% 
  filter(n %in% c(364, 104, 416))
```

It seems like these combinations are not in the raw data. I check the annotation table.

```{r}
read_csv(here("anno/combi_validation_robot.csv")) %>% 
  separate(drug_pair_conc, c("drug_1", "drug_2", "conc_1", "conc_2"), 
           remove = FALSE, sep = "_") %>%
  mutate(block_id = paste(drug, sep = "_")) %>% 
  group_by(block_id) %>% 
  summarise(n = n())
```


## synergy calculation

Now having this wrangled dataframe and an own function in place, I calculated synergies by calling the "calculate_synergy.R" script. 

After the values have been calculated I go on and check the distribution of scores.

```{r}
order <- syn_scores %>% 
  filter(line != "D093N11") %>% 
  group_by(block_id) %>% 
  summarise(mean = mean(score_average)) %>% 
  arrange(mean) %>% 
  .$block_id

syn_scores %>% 
  filter(line != "D093N11") %>%
  mutate(block_id = factor(block_id, levels = order)) %>%
  ggplot(aes(score_average, block_id, color = line)) + 
  geom_vline(xintercept = 0) +
  geom_jitter(height = 0.1, width = 0.1) +
  facet_wrap(~ method ) + 
  theme_cowplot() + 
  scale_colour_brewer(type = "qual", palette = 3)
```

I wonder it synergy scores correlated well between replicates. 

```{r}
r <- syn_scores %>%
  filter(line != "D093N11") %>%
  mutate(id = paste(block_id, line, method, sep = "_")) %>%
  dplyr::select(replicate, score_average, id) %>%
  tidyr::spread(replicate, score_average) %>% 
  rename(rep1 = `1`) %>%
  rename(rep2 = `2`) %>%
  dplyr::select(rep1, rep2) %>% 
  as.matrix() %>%
  cor(method = "spearman") %>%
  .["rep1", "rep2"]

df <- syn_scores %>% 
  filter(line != "D093N11") %>%
  mutate(id = paste(block_id, line, method, sep = "_")) %>%
  dplyr::select(replicate, score_average, id) %>%
  tidyr::spread(replicate, score_average) %>% 
  rename(rep1 = `1`) %>%
  rename(rep2 = `2`) 

df %>% write_csv(here::here("reports/plot_data/correlation_synergy.csv"))

df %>%
  ggplot(aes(rep1, rep2)) + 
  geom_point(alpha = 0.2) +
  #geom_point(alpha = 0.2, aes(color = ctrl)) + 
  #scale_color_manual(values=c("#E69F00", "#999999", "#56B4E9")) +
  geom_abline(intercept = 0, slope = 1) +
  #geom_hline(yintercept = 1) +
  theme_classic() + 
  ggtitle(paste0("r (spearman) = ", r %>% round(2))) + 
  xlab("replicate 1") + 
  ylab("replicate 2") + 
  coord_fixed() + 
  ggsave(here::here("reports/figures/correlation_synergy.pdf"), width = 4, height = 4)
```

```{r}
order_line <- syn_scores %>% 
  filter(line != "D093N11") %>% 
  group_by(line) %>% 
  summarise(mean = mean(score_average)) %>% 
  arrange(mean) %>% 
  .$line

syn_scores %>% 
  filter(line != "D093N11") %>%
  mutate(line = factor(line, levels = order_line)) %>%
  ggplot(aes(score_average, line, color = block_id)) + 
  geom_vline(xintercept = 0) +
  geom_jitter(height = 0.1, width = 0.1) +
  facet_wrap(~ method ) + 
  theme_cowplot() + 
  scale_colour_brewer(type = "qual", palette = 3)
  
```




The correlation is okayish
I decide to fit a linear model of both replicate and block_id on the data

```{r}
syn_collapsed <- syn_scores %>% 
  filter(line != "D093N11") %>%
  filter(block_id != "Vorinostat_EHMT-Inh.") %>%
  group_by(block_id, method, line) %>% 
  summarise(score = mean(score_average))

syn_collapsed %>% write_csv(here("data/processed/synergy_simplified.csv"))

# lm(score ~ block_id, data = syn_collapsed %>% filter(method == "ZIP")) %>% summary()
# lm(score ~ block_id, data = syn_collapsed %>% filter(method == "Bliss")) %>% summary()
# lm(score ~ block_id, data = syn_collapsed %>% filter(method == "Loewe")) %>% summary()
# lm(score ~ block_id, data = syn_collapsed %>% filter(method == "HSA")) %>% summary()

lm(score ~ block_id + method, data = syn_collapsed) %>% summary()
```


I wonder if certain subtypes are becoming obvious

```{r}
syn_collapsed %>% 
  ungroup() %>%
  filter(method == "ZIP") %>% 
  dplyr::select(line, score, block_id, -method) %>% 
  
  spread(line, score) %>% 
  as.data.frame() %>% 
  column_to_rownames("block_id") %>% 
  pheatmap::pheatmap(filename = here::here("reports/figures/heatmap_zip_line.pdf"))
```


```{r}
set.seed(1234)

model_estimate <- lm(score ~ block_id + method, data = syn_collapsed %>% filter(line != "D093N11")) %>% broom::tidy() %>% 
  mutate(q = p.adjust(p.value, method = "bonferroni")) %>% 
  filter(q < 0.05, grepl(pattern = "block", x = term)) %>% 
  mutate(term = substr(term, 9, nchar(term))) %>% 
  arrange(desc(estimate))

model_estimate %>% write_csv(here::here("reports/tables/lm_synergy.csv"))

model_estimate
```


```{r}
syn_collapsed %>%
  ungroup() %>%
filter(line != "D093N11") %>%
  mutate(block_id = factor(block_id, levels = order)) %>%
  ggplot(aes(score, block_id, color = line)) + 
  geom_vline(xintercept = 0) +
  geom_jitter(height = 0.1, width = 0.1) +
  facet_wrap(~ method ) + 
  theme_cowplot() + 
  scale_colour_brewer(type = "qual", palette = 3) + 
  # coord_fixed() +
  ggsave(here::here("reports/figures/synergy_line.pdf"))
```
