---
title: "01_inspect_dataset"
author: "Niklas Rindtorff"
output: 
   BiocStyle::html_document2:
    code_folding: hide
    df_print: paged
    toc_depth: 3
    toc_float: true
    fig_height: 6
    fig_width: 9.5
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
# Set some global knitr options
library("knitr")
opts_chunk$set(tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=60), cache=TRUE, messages=FALSE)
```

#Load packages

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(pheatmap)
library(splots)
library(patchwork)
library(synergyfinder)
library(ggsignif)
library(broom)
library(ggrepel)
```

#Load data

```{r, warning=FALSE, message=FALSE}
##read ctg data files
ctg_data <- lapply(list.files('~/tas/data/ctg_data/', full.names=T, pattern = "TAS"), 
                   function(f) read_tsv(f, col_names=F, col_types = 'cci') %>%
                     `colnames<-`(c('screen', 'well', 'pcount'))) %>% bind_rows() %>%
  separate(screen, c("date", "operator", "mithras", "experiment_id"), sep = "_", remove = FALSE) %>%
  separate(experiment_id, c("experiment", "line", "TAS_concentration")) %>%
  #join compound annotation data
  full_join(., lapply(list.files('~/tas/data/annotations/', full.names=T), 
                   function(f) read_csv(f)) %>% bind_rows() %>%
  rename(well = destination.well) %>%
  select(drug, well, Source_Position, Source) %>%
  separate(well, c("row", "col"), sep = 1, remove = FALSE) %>%
  mutate(drug = if_else(is.na(drug), Source, drug),
         drug = if_else(well == "J19", "DMSO", drug),
         combination = if_else(col == "13", FALSE, TRUE))) %>%
  unite(id, c("date", "line", "TAS_concentration", "experiment"), remove = FALSE) %>%
  mutate(row_num = match(row, LETTERS[1:26]),
         col_num = as.numeric(col)) %>%
  #annotate concentrations
  mutate(concentration = ifelse(Source_Position == "Rack", NA,
                                ifelse(Source_Position == "Master_Library_Plate", 1,
                                       ifelse(Source_Position == "Dilution_Plate_1", 2,
                                              ifelse(Source_Position == "Dilution_Plate_2", 3,
                                                     ifelse(Source_Position == "Dilution_Plate_3", 4,
                                                            ifelse(Source_Position == "Dilution_Plate_4", 5, "no match")))))))
```


#Inspect and clean screening data

On the second screening day some plates were read-out twice. They are labeled TAS and TASa. I check if there are profound differences between these plates. 
First I correlate the pcount of these plates. 

```{r}
ctg_data %>% 
  filter(date == "171212" & line == "19") %>%
  select(pcount, well, id) %>%
  spread(id, pcount) %>%
  select(-well) %>%
  as.data.frame() %>%
  cor() %>%
  pheatmap()
```

The plates are virtually the same. I also checked differences between zfactors, which are not apparent. I keep the plates with the experiment name "TAS" for simplicity. 

```{r}
ctg_data <- ctg_data %>% filter(experiment != "TASa")
```

Now I want to gain a broad overview of the asssay plates.

```{r, warning=FALSE}
ctg_data %>% 
  select(id, pcount) %>% split(., .$id) %>% lapply(., function(f) f$pcount %>% as.vector) %>%
  plotScreen(., 6)
```

Now I focus on column and row-wise effects. Row wise effects dominate clearly

```{r}
ctg_data %>% ggplot(aes(row, pcount)) + geom_point() + 
  geom_smooth() +
  facet_wrap(~id) + theme_bw() + ggtitle("Spatial effects of rows") + 
  #next figure
  ctg_data %>% ggplot(aes(col, pcount)) + geom_point() + 
  geom_smooth() +
  facet_wrap(~id) + theme_bw() + ggtitle("Spatial effects of columns")
```

I see some spatial effects on most screening plates. The effect is mostly restricted to row-wise differences. I apply a conservative loss smoothing to correct for these.

```{r, message=FALSE}
## split data by plate and apply loess normalization
ctg_loess <- ctg_data %>% dplyr::select(row_num, col_num, pcount, id) %>% 
  # na values
  drop_na() %>%
  #set the ctrl column 13 to NA
  #mutate(pcount = ifelse(col_num == 13, NA, pcount)) %>%
  split(.$id) %>% lapply(function(s){
  ## loess fit. family is 'symmetric' to be robust to outliers
  fit <- loess(pcount ~ row_num + col_num, data=s, family='symmetric')
  ## apply normalization
  tibble(norm_fac = fit$fitted) %>% cbind(s %>% drop_na(),.) %>% 
    mutate(pcount_norm = pcount - (norm_fac - median(norm_fac)))
}) %>% bind_rows() %>% full_join(ctg_data)
```

Now I plot the results. 

```{r}
ctg_loess %>% ggplot(aes(row, pcount_norm)) + geom_point() + 
  geom_smooth() +
  facet_wrap(~id) + theme_bw() + ggtitle("Spatial effects of rows") + 
  #next figure
  ctg_loess %>% ggplot(aes(col, pcount_norm)) + geom_point() + 
  geom_smooth() +
  facet_wrap(~id) + theme_bw() + ggtitle("Spatial effects of columns")
```

For demo purposes I show one plate before and after correction: 

```{r}
ctg_data %>% 
  filter(TAS_concentration == "1") %>% 
  ggplot(aes(row, pcount)) + geom_point() + 
  #geom_smooth() +
  facet_wrap(~line) + theme_bw() + ggtitle("Before Normalization: Spatial effects of rows") + 
  #and after normalization
  ctg_loess %>% 
  filter(TAS_concentration == "1") %>% 
  ggplot(aes(row, pcount_norm)) + geom_point() + 
  geom_smooth() +
  facet_wrap(~line) + theme_bw() + ggtitle("After Normalization:Spatial effects of rows")
```


#QC
Now I look at the distribution of controls in the first screening replicate.

```{r, warning=FALSE}
coi <- c("Stauro_500nM", "DMSO")

ctg_loess %>%
  mutate(line_date_exp = paste0(line, "_", date, "_", experiment)) %>%
  filter(drug %in% coi) %>%
  ggplot(aes(drug, pcount_norm, color = combination)) +
  geom_jitter(alpha = 0.7) + 
  facet_grid(line_date_exp ~ TAS_concentration) + 
  theme_bw()
```

The distribution seems to be quiet good. I now go on calculating the z-factors for each plate for combined and uncombined wells. 

```{r, warning=FALSE}
ctg_loess %>%
  filter(drug %in% coi) %>%
  group_by(id, drug, combination, TAS_concentration, line, date, experiment) %>%
  summarise(sd = sd(pcount_norm, na.rm = TRUE),
            mean = mean(pcount_norm, na.rm = TRUE)) %>%
  group_by(id, combination, TAS_concentration, line, date, experiment) %>%
  summarise(zfactor = 1-((3*sum(sd))/abs(range(mean)[1]-range(mean)[2]))) %>%
  mutate(qc = if_else(zfactor < 0.25, FALSE, TRUE)) %>%
  ggplot(aes(combination, zfactor, color = qc, shape = date)) +
  geom_point(alpha = 0.7, size = 5) + 
  facet_grid(line ~ TAS_concentration) + 
  theme_bw() + 
  geom_hline(yintercept = 0.25)
```

Also the z-factors are okay for most plates. Three plates have slight losses in assay quality, however, this seems still acceptable.

#Normalize screening data

Again, I take a look at the TAS-untreated ctrls. They should be the same across plates

```{r}
coi <- c("DMSO", "Stauro_500nM")

ctg_loess %>%
  filter(drug %in% coi & combination == FALSE) %>%
  ggplot(aes(TAS_concentration, pcount_norm)) +
  geom_boxplot(aes(colour = drug), width = 6) + 
  theme_bw() + 
  facet_grid(line~date) +
  #scale_y_continuous(limits = c(-1,250000)) +
  ggtitle("Before DMSO normalization") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

There are some line specific differences in viability and one outlier plate 171128_27_4. I now normalize the plates to the untreated DMSO control.

```{r}
ctg_ln <- ctg_loess %>%
  full_join(., ctg_loess %>% filter(drug == "DMSO" & combination == FALSE) %>% group_by(id) %>% 
              summarise(ctrl = median(pcount_norm, na.rm = TRUE)) %>% ungroup()) %>%
  group_by(id) %>% mutate(rv = pcount_norm/ctrl) %>% ungroup() 
```

Now I look at the controls after normalization. 

```{r}
coi <- c("DMSO", "Stauro_500nM")

ctg_ln %>%
  filter(drug %in% coi & combination == FALSE) %>%
  ggplot(aes(TAS_concentration, rv)) +
  geom_boxplot(aes(colour = drug), width = 6) + 
  theme_bw() + 
  facet_grid(line~date) +
  #scale_y_continuous(limits = c(-1,250000)) +
  ggtitle("After DMSO normalization") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

The normalization worked well. As intended there are controlled line and TAS dependent effects left. 

```{r}
ctg_ln %>%
  ggplot(aes(TAS_concentration, rv)) +
  geom_boxplot( width = 6) + 
  theme_bw() + 
  facet_grid(line~date) +
  #scale_y_continuous(limits = c(-1,250000)) +
  ggtitle("Relative viability measurments for every plate") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

I wonder if I have to center the dataset further inbetween 1 and 0. Therefore I plot a histogram of all viabilities.

```{r}
ctg_ln %>%
  ggplot(aes(rv)) + 
  geom_histogram(bins = 90) + 
  theme_classic() + 
  geom_vline(xintercept = c(0,1)) + 
  xlab("relative viability")
```

As an alternative approach I trim the data between 1 and 0, assuming that these datapoints are only spread and normalization artifacts. 

```{r}
ctg_lnt <- ctg_ln %>% 
  mutate(
    rv_trim = ifelse(rv < 0, 0, rv),
    rv_trim = ifelse(rv > 1, 1, rv_trim)
         )

ctg_lnt %>%
  ggplot(aes(rv_trim)) + 
  geom_histogram(bins = 90) + 
  theme_classic() + 
  geom_vline(xintercept = c(0,1)) + 
  xlab("relative viability")
```

I am fine with this distribution. Let's move on. 
Before we continue I take a look at the overall correlation. 

```{r}
coi <- c("DMSO", "Stauro_500nM")

tmp <- ctg_ln %>%
  unite("line_well", c("line", "well", "TAS_concentration")) %>%
  ungroup() %>%
  mutate(Replicate = ifelse(date == "171128", 1, 2)) %>%
  dplyr::select(line_well, Replicate, rv, drug) %>%
  spread(Replicate, rv) %>%
  mutate(ctrl = if_else(drug %in% coi, TRUE, FALSE))

tmp.cor<- tmp %>% 
  dplyr::select(`1`,`2`) %>%
  cor(., use = "pairwise.complete.obs") %>%
  .[1,2] %>%
  round(2)

tmp %>%
  ggplot(aes(x = `1`, y = `2`)) + 
  geom_point(alpha = 0.3) + 
  theme_classic() + 
  ylab("Replicate 2") + 
  xlab("Replicate 1") + 
  ggtitle(paste0("correlation of normalized viabiliy, r =", tmp.cor)) + 
  geom_abline(slope = 1) + 
  #geom_density2d(alpha = 0.7, color = "black") + 
  scale_x_continuous(breaks = c(0, 0.5, 1)) + 
  scale_y_continuous(breaks = c(0, 0.5, 1)) +
  geom_hline(yintercept = 1) + 
  geom_vline(xintercept = 1)
  #stat_binhex() 
```


#Synergy determination

I consider using the "synergyfinder" package, published by He et al. at the FIMM. 
First I have to wrangle the data into a standardized input format. For this I write an ugly function. 

```{r}
prepare_synergy <- function(df, controls){
  #not a universal function
  syn_drug <- df %>%
    filter(!drug %in% controls & combination == TRUE) %>%
    #setup the easy stuff
    mutate(BlockID = paste("TAS-102", drug, sep = "_"),
           Response = rv*100,
           Replicate = ifelse(date == "171128", 1, 2),
           DrugRow = "TAS-102",
           DrugCol = drug,
           #build matrix
           Col = ifelse(drug != "DMSO", as.numeric(concentration), 6),
           Row = ifelse(TAS_concentration != "0", as.numeric(TAS_concentration), 6),
           #define concentrations
           ConcRow = ifelse(TAS_concentration != "0", 25/(5^(as.numeric(TAS_concentration)-1)), 0), #assuming the initila concentration is 50
           ConcCol = 10/(5^(as.numeric(concentration)-1)),
           ConcUnit = "uM"
           )
  
  syn_ctrl <- full_join(syn_drug %>% select(BlockID, line, TAS_concentration, Replicate) %>% distinct(),
                       #add TAS DMSO controls
                     df %>%
              filter(drug == "DMSO" & combination == TRUE) %>%
              mutate(Replicate = ifelse(date == "171128", 1, 2)) %>%
              group_by(line, TAS_concentration, Replicate) %>% 
              mutate(Response = median(rv, na.rm = TRUE)*100) %>%
              mutate(#define concentrations
                     ConcRow = ifelse(TAS_concentration != "0", 50/(5^(as.numeric(TAS_concentration)-1)), 0), #assuming the initila concentration is 50
                     ConcCol = 0,
                     Col = 6,
                     Row = ifelse(TAS_concentration != "0", as.numeric(TAS_concentration), 6),
                     DrugRow = "TAS-102",
                     ConcUnit = "uM"
                       )) %>%
                    #hacky way of formatting the data
                mutate(DrugCol = substr(BlockID, 9, nchar(BlockID)))
  
  syn_mat <- full_join(syn_drug, syn_ctrl) %>% dplyr::select(BlockID, Col, Row, Response, Replicate, DrugRow, DrugCol, ConcRow, ConcCol, ConcUnit, line) %>% distinct()
  
  return(syn_mat)
}
```


Now I apply this function. 

```{r, warning=FALSE}
coi <- c("DMSO", "Stauro_500nM")
syn_mat <- prepare_synergy(ctg_lnt, coi)

#I also determine the synergy for the trimmed dataset
syn_mat_t <- prepare_synergy(ctg_lnt %>% dplyr::select(-rv) %>% rename(rv = rv_trim), coi)
```

Now I define a wrapper function that can apply multiple synergy calculations and report the results in a tidy format. 
For now I stick to calulating the average synergy scores.

```{r, eval = TRUE}
tidy_synergy = function(df, me){ lapply(as.list(me), function(m, f = df){
  #hacky way of ignoring cases in which the synergy scores could not be calculated
  return(tryCatch(f %>% 
      ReshapeData(., data.type = "viability") %>%
      #PlotDoseResponse() %>%
      CalculateSynergy(method = m, correction = TRUE) %>%
    tibble(BlockID = .$drug.pairs %>% simplify() %>% .[4],
           method = .$method,
           score_average = .$scores %>% .[[1]] %>% .[-1,-1] %>% simplify() %>% mean(),
           score_sd = .$scores %>% .[[1]] %>% .[-1,-1] %>% simplify() %>% sd()) %>%
    select(BlockID, method, score_average, score_sd) %>% distinct(), 
    error=function(e) NULL))}
) %>% bind_rows()
  #PlotSynergy(type = "all")
  }
```

I apply the function on the dataset and save/load it. 

```{r, warning=FALSE, message=FALSE, error=FALSE, eval = FALSE}
tas_syn <- syn_mat %>%  group_by(line, Replicate , BlockID) %>% 
  do(tidy_synergy(df = ., me = c("ZIP", "HSA", "Bliss", "Loewe"))) %>%
  ungroup() %>%
  #complete the df for scores that could not be calculated
  mutate(method = as.factor(method),
         line = as.factor(line)) %>% 
  complete(BlockID, method, Replicate, line) %>%
  mutate(score_average = ifelse(is.nan(score_average), NA, score_average))


save(tas_syn, file = "tas_syn_complete.Rdata")
```

```{r}
load("tas_syn_complete.Rdata")
```


As an alternative I calculate synergy scores only for selected query concentrations.
I start out by skipping the highest tested (TAS) concentration. 

```{r, warning=FALSE, message=FALSE, error=FALSE}
syn_mat_low <- syn_mat %>% filter(Row != 1 ) %>% mutate(Row = Row -1)

# syn_mat_low <- syn_mat %>% filter(Row != 1 & Col != 1) %>% mutate(Row = Row -1,
#                                                                   Col = Col -1)

tas_syn_l <- syn_mat_low %>%  group_by(line, Replicate , BlockID) %>%
  do(tidy_synergy(df = ., me = c("ZIP", "HSA", "Bliss", "Loewe"))) %>%
  ungroup() %>%
  #complete the df for scores that could not be calculated
  mutate(method = as.factor(method),
         line = as.factor(line)) %>%
  complete(BlockID, method, Replicate, line) %>%
  mutate(score_average = ifelse(is.nan(score_average), NA, score_average))

save(tas_syn_l, file = "tas_syn_low.Rdata")
```

```{r}
load("tas_syn_low.Rdata")
```

##Comparing dropout synergy calculation

First I wonder how many models were fitted in both approaches.

```{r}
df <- tibble(complete = tas_syn$score_average,
       wo_highest = tas_syn_l$score_average)

tibble(complete_na = df$complete %>% is.na() %>% sum(),
       wo_highest_na = df$wo_highest %>% is.na() %>% sum(),
       frac_comp = complete_na/nrow(df),
       frac_wo_hi = wo_highest_na/nrow(df))
```



Now I am interested in the correltaion between synergy scores 

```{r}
tibble(complete = tas_syn$score_average,
       wo_highest = tas_syn_l$score_average) %>%
  ggplot(aes(complete, wo_highest)) + 
  geom_point() + 
  theme_classic()

tibble(complete = tas_syn$score_average,
       wo_highest = tas_syn_l$score_average) %>%
  corrr::correlate()
```

For now I add the synergy data which was calculated without the highest concentration to the dataset.

```{r}
tas_syn <- full_join(tas_syn, tas_syn_l %>% 
                       rename(score_average_l = score_average,
                              score_sd_l = score_sd))
                       
```


Now I want to inspect the distribution of synergy scores.

```{r}
tas_syn %>% 
  ggplot(aes(score_average)) +
  geom_histogram() +
  facet_grid(line ~ method)
```

For now, I will stick to to ZIP method which calculates a delta-score. In recent publications this score is >15 for strong interactions and <-10 for antagonistic combinations.
I observed that on multiple occasions not all statistics could be calculated. I want to see how often theses errors happen. 

```{r}
# tas_syn %>% 
#   group_by(BlockID, method) %>%
#   summarise(n = sum(is.na(score_average))) %>%
#   filter(n != 0) %>%
#   ggplot(aes(n)) +
#   geom_histogram() +
#   facet_grid(~ method)

#define list
fb <- tas_syn %>%
  group_by(BlockID, method) %>%
  summarise(n = sum(is.na(score_average))) %>%
  filter(n > 4) %>%
  .$BlockID
```

Errors during calculation happen in about 10% of all drug combinations, probably due to an inability to a model for the data. I define a list of combinations which will be excluded for now due to too sparse data.

```{r}
tas_test <- lapply(as.list(tas_syn %>% filter(!BlockID %in% fb) %>%. $BlockID %>% unique()), 
                   function(boi) tas_syn %>%
                     filter(!BlockID %in% fb) %>%
                     mutate(test = if_else(BlockID == boi, TRUE, FALSE),
                            f_method = method) %>% 
                     group_by(f_method) %>%
                     do(tidy(t.test(score_average~test, data=.))) %>% 
                     mutate(BlockID = boi)) %>% 
  bind_rows() %>%
  group_by(f_method) %>%
  mutate(p.adj = p.value %>% p.adjust(method = "BH"))

tas_test %>% 
  ggplot(aes(p.adj)) +
  geom_histogram() + 
  facet_wrap(~f_method) + 
  theme_bw()
```

#Synergy plotting

##Synergy Dotplots
Now I am interested in compounds which are significantly different from the rest of all tested combinations.

```{r, eval = FALSE}
tas_test %>%
  group_by(BlockID) %>%
  mutate(avg_p = mean(p.adj, na.rm = TRUE)) %>%
  arrange(avg_p) %>%
  ungroup() %>%
  mutate(BlockID = factor(BlockID, levels = BlockID %>% unique())) %>%
  ggplot(aes(BlockID, p.adj)) +
  geom_point() + 
  theme_minimal() + 
  scale_y_log10() + 
  facet_wrap(~f_method) + 
  geom_hline(yintercept = 0.05) + 
  geom_label_repel(data = tas_test %>% filter(p.adj < 0.05), aes(label = BlockID)) + 
  theme(
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```

Next to one combination with strong synergism I also look for other potent combinations. 
In the following plot I show the synergy scores for all categories and in red the synergy score based on the drop-out data.

```{r}
tas_syn %>%
group_by(BlockID, method) %>%
  mutate(avg_s = mean(score_average, na.rm = TRUE),
         avg_sl = mean(score_average_l, na.rm = TRUE)) %>%
  group_by(BlockID) %>%
  mutate(avg_a = mean(score_average, na.rm = TRUE)) %>%
  arrange(avg_a) %>%
  ungroup() %>%
  mutate(BlockID = factor(BlockID, levels = BlockID %>% unique())) %>%
  ggplot(aes(BlockID, avg_s)) +
  geom_point(color = "black") + 
  geom_point(aes(BlockID, avg_sl), color = "red") + 
  theme_classic() + 
  geom_hline(yintercept = 0) + 
  #scale_y_log10() + 
  facet_wrap(~method) + 
  #geom_hline(yintercept = 0.05) + 
  geom_label_repel(data = tas_syn %>% group_by(BlockID, method) %>%
    summarise(avg_s = mean(score_average, na.rm = TRUE)) %>% 
    group_by(method) %>%
      do(arrange(., desc(avg_s)) %>% head(3)), aes(label = BlockID), hjust = -0.1, vjust = -0.1) + 
  ylab("Average Synergy Score") + 
  xlab("Drug Combinations") + 
  theme(
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```


Now I invert the color scheme and arrange compounds according to the drop-out synergies. In blue I show the "complete" synergy scores.

```{r}
tas_syn %>%
group_by(BlockID, method) %>%
  mutate(avg_s = mean(score_average, na.rm = TRUE),
         avg_sl = mean(score_average_l, na.rm = TRUE)) %>%
  group_by(BlockID) %>%
  mutate(avg_al = mean(score_average_l, na.rm = TRUE)) %>%
  arrange(avg_al) %>%
  ungroup() %>%
  mutate(BlockID = factor(BlockID, levels = BlockID %>% unique())) %>%
  ggplot(aes(BlockID, avg_sl)) +
  geom_point(color = "black") + 
  geom_point(aes(BlockID, avg_s), color = "blue") + 
  theme_classic() + 
  geom_hline(yintercept = 0) + 
  #scale_y_log10() + 
  facet_wrap(~method) + 
  #geom_hline(yintercept = 0.05) + 
  geom_label_repel(data = tas_syn %>% group_by(BlockID, method) %>%
    summarise(avg_sl = mean(score_average_l, na.rm = TRUE)) %>% 
    group_by(method) %>%
      do(arrange(., desc(avg_sl)) %>% head(3)), aes(label = BlockID), hjust = -0.1, vjust = -0.1) + 
  ylab("Average Synergy Score") + 
  xlab("Drug Combinations") + 
  theme(
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```

The Bliss score is an established measure of drug synergy. Therefore I plot it in more detail.
First I plot only the complete averages across all lines and replicates.

```{r}
tas_syn %>%
  filter(method == "Bliss") %>%
group_by(BlockID, method) %>%
  mutate(avg_s = mean(score_average, na.rm = TRUE)) %>%
  group_by(BlockID) %>%
  mutate(avg_a = mean(score_average, na.rm = TRUE)) %>%
  arrange(avg_a) %>%
  ungroup() %>%
  mutate(BlockID = factor(BlockID, levels = BlockID %>% unique())) %>%
  ggplot(aes(BlockID, avg_s)) +
  geom_point() + 
  theme_classic() + 
  geom_hline(yintercept = 0) + 
  #scale_y_log10() + 
  #geom_hline(yintercept = 0.05) + 
  geom_label_repel(data = tas_syn %>% filter(method == "Bliss") %>% group_by(BlockID, method) %>%
    summarise(avg_s = mean(score_average, na.rm = TRUE)) %>% 
    group_by(method) %>%
      do(arrange(., desc(avg_s)) %>% head(5)), aes(label = BlockID), hjust = -0.1, vjust = -0.1) + 
  ylab("Bliss Synergy Score") + 
  xlab("Drug Combinations") + 
  theme(
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```

## Synergy Boxplot
To see if there are line specific differences, I plot the data in a boxplot format. The first plot shows the complete synergy dataset.

```{r}
tas_syn %>%
  filter(method == "ZIP") %>%
group_by(BlockID, method) %>%
  mutate(avg_s = mean(score_average, na.rm = TRUE),
         avg_sl = mean(score_average_l, na.rm = TRUE)) %>%
  group_by(BlockID) %>%
  mutate(avg_a = mean(score_average, na.rm = TRUE)) %>%
  arrange(avg_a) %>%
  ungroup() %>%
  mutate(BlockID = factor(BlockID, levels = BlockID %>% unique())) %>%
  ggplot(aes(BlockID, score_average)) +
  geom_boxplot() +
  geom_point(aes(color = line)) + 
  theme_classic() + 
  geom_hline(yintercept = 0) + 
  #scale_y_log10() + 
  #geom_hline(yintercept = 0.05) + 
  ylab("ZIP Synergy Score") + 
  xlab("Drug Combinations") + 
  theme(
        axis.text.x = element_text(angle = 90, hjust = 1),
        axis.ticks.x=element_blank())
```

Now I take a look at the "dropout" dataset in a similiar way.

```{r}
tas_syn %>%
  filter(method == "ZIP") %>%
group_by(BlockID, method) %>%
  mutate(avg_s = mean(score_average, na.rm = TRUE),
         avg_sl = mean(score_average_l, na.rm = TRUE)) %>%
  group_by(BlockID) %>%
  mutate(avg_al = mean(score_average_l, na.rm = TRUE)) %>%
  arrange(avg_al) %>%
  ungroup() %>%
  mutate(BlockID = factor(BlockID, levels = BlockID %>% unique())) %>%
  ggplot(aes(BlockID, score_average_l)) +
  geom_boxplot() +
  geom_point(aes(color = line)) + 
  theme_classic() + 
  geom_hline(yintercept = 0) + 
  #scale_y_log10() + 
  #geom_hline(yintercept = 0.05) + 
  ylab("ZIP Synergy Score") + 
  xlab("Drug Combinations") + 
  theme(
        axis.text.x = element_text(angle = 90, hjust = 1),
        axis.ticks.x=element_blank())
```

##Synergy Heatmaps

```{r, eval = FALSE}
tas_syn %>% 
  unite(exp ,c(method, Replicate, line)) %>%
  dplyr::select(exp, BlockID, score_average, BlockID) %>%
  as.data.frame() %>%
  spread(exp, score_average) %>%
  remove_rownames() %>%
  column_to_rownames(., var = "BlockID") %>%
  drop_na() %>%
  pheatmap()
  
```

#Downsampling of query concnetrations

##Build sampler

I have 5 concentrations plus DMSO control. I will only measure the ZIP score later on, as this score is more robust

```{r}
#I create a list for all possible combinations
combn(1:5, 4) %>% split(., rep(1:ncol(.), each = nrow(.))) %>% 
  append(., combn(1:5, 3) %>% split(., rep(1:ncol(.), each = nrow(.)))) %>% 
  append(., combn(1:5, 2) %>% split(., rep(1:ncol(.), each = nrow(.)))) %>%
  append(., combn(1:5, 5) %>% split(., rep(1:ncol(.), each = nrow(.)))) %>%
  append(., combn(1:5, 1) %>% split(., rep(1:ncol(.), each = nrow(.)))) -> combinations

names(combinations) <- combinations %>% lapply(.,paste, collapse = "") %>% unlist

```

Now I build a selector that takes all the desired concentrations and drops others

```{r}
ds_selector <- function(n){
  syn_mat_f <- syn_mat %>% filter(Row %in% n) 
  
  if(length(n) == 1){
    syn_mat_fr <- syn_mat_f %>% 
      mutate(Row = replace(Row, Row == n[1], 1))
  }
  if(length(n) == 2){
    syn_mat_fr <- syn_mat_f %>% 
      mutate(Row = replace(Row, Row == n[1], 1)) %>%
      mutate(Row = replace(Row, Row == n[2], 2)) 
  }
  if(length(n) == 3){
    syn_mat_fr <- syn_mat_f %>% 
      mutate(Row = replace(Row, Row == n[1], 1)) %>%
      mutate(Row = replace(Row, Row == n[2], 2)) %>%
      mutate(Row = replace(Row, Row == n[3], 3))
  }
  if(length(n) == 4){
    syn_mat_fr <- syn_mat_f %>% 
      mutate(Row = replace(Row, Row == n[1], 1)) %>%
      mutate(Row = replace(Row, Row == n[2], 2)) %>%
      mutate(Row = replace(Row, Row == n[3], 3)) %>%
      mutate(Row = replace(Row, Row == n[4], 4))
  }
  
  syn_mat_fr <- syn_mat_fr %>% mutate(selector = paste(n, collapse = ""))
  return(syn_mat_fr)
}
```

Now I run the selector and calculate synergy scores. I save and reload the data

```{r, warning = FALSE, message = FALSE, eval = FALSE}

tas_syn_sampling <- lapply(combinations, ds_selector)  %>% lapply(., function(df) df %>% group_by(line, Replicate , BlockID, selector) %>%
  do(tidy_synergy(df = ., me = c("ZIP")))) %>% 
  bind_rows() %>%
  ungroup() %>%
  #complete the df for scores that could not be calculated
  mutate(method = as.factor(method),
         line = as.factor(line)) %>%
  complete(BlockID, method, Replicate, line) %>%
  mutate(score_average = ifelse(is.nan(score_average), NA, score_average))

save(tas_syn_sampling, file = "tas_syn_sampling.Rdata")
```

```{r}
load("tas_syn_sampling.Rdata")
```

```{r}
tas_syn_sampling %>% 
  dplyr::select(-score_sd) %>% 
  unite(id, BlockID, method, Replicate, line) %>%
  spread(selector, score_average) %>% 
  as.data.frame() %>% 
  remove_rownames() %>% 
  column_to_rownames("id") %>% 
  cor(use = "pairwise.complete.obs") %>% 
  as.data.frame() %>%
  .[colSums(!is.na(.)) > 0] %>%
  drop_na() %>% 
  pheatmap()
```

I rank the different combinations by their correlation with the complete dataset's synergy scores

```{r}
tas_syn_sampling %>% 
  dplyr::select(-score_sd) %>% 
  unite(id, BlockID, method, Replicate, line) %>%
  spread(selector, score_average) %>% 
  as.data.frame() %>% 
  remove_rownames() %>% 
  column_to_rownames("id") %>% 
  cor(use = "pairwise.complete.obs", method = "spearman") %>% 
  as.data.frame() %>%
  rownames_to_column("id") %>%
  as_tibble() %>% 
  dplyr::select(id, "12345") %>% 
  rename(r = "12345") %>% 
  mutate(n = nchar(id) %>% as.character()) %>%
  left_join(., tas_syn_sampling %>% 
  dplyr::select(-score_sd) %>% 
  unite(id, BlockID, method, Replicate, line) %>%
  complete(id, selector) %>%
  group_by(selector) %>% 
  mutate(na = score_average %>% is.na() %>% sum()) %>% dplyr::select(selector, na) %>% rename(id = selector) %>% distinct()) %>%
  filter(n != 5) %>% 
  arrange((r)) %>% 
  mutate(id = factor(id, levels = id)) %>% 
  
  ggplot(aes(id, r)) + 
  geom_point(aes(color = n, size = na)) + 
  theme_classic() + 
  ggtitle("Correlation of downsampled synergy scores for selected concentrations (1 is highest, 5 is lowest)") + 
  ylab("Spearman r") + 
  xlab("Combination of concentrations (size corresponds to cases where no synergy could be calculated, color to number of chosen concentrations)")
  
```


#Examples
Now I take a closer look at some drug combinations

```{r}
boi = c("TAS-102_MK-8776/ SCH900776")
moi = "ZIP"

tas_syn %>%
  mutate(boi = if_else(BlockID %in% boi, TRUE, FALSE)) %>% 
  filter(method == moi) %>%
  ggplot(aes(as.factor(boi), score_average, group=boi)) +
  geom_jitter(aes(color=line), width = 0.3) + 
  stat_summary(fun.y = 'mean', fun.ymin ='mean', fun.ymax='mean', colour = 'black', geom='crossbar', width=0.5) + 
  geom_signif(comparisons = list(c("TRUE", "FALSE")), method = "t.test") + 
  theme_classic() + 
  ggtitle(paste0(boi," method:", moi))

syn_mat %>% filter(line == "07", BlockID == boi, Replicate == 1) %>%
  ReshapeData(., data.type = "viability") %>%
  CalculateSynergy(method = "ZIP", correction = TRUE) %>%
  PlotSynergy(type = "all", save.file = FALSE)
```

```{r}
boi = c("TAS-102_Birinapant")
moi = "ZIP"

tas_syn %>%
  mutate(boi = if_else(BlockID %in% boi, TRUE, FALSE)) %>% 
  filter(method == moi) %>%
  ggplot(aes(as.factor(boi), score_average, group=boi)) +
  geom_jitter(aes(color=line), width = 0.3) + 
  stat_summary(fun.y = 'mean', fun.ymin ='mean', fun.ymax='mean', colour = 'black', geom='crossbar', width=0.5) + 
  geom_signif(comparisons = list(c("TRUE", "FALSE")), method = "t.test") + 
  theme_classic() + 
  ggtitle(paste0(boi," method:", moi))

syn_mat %>% filter(line == "07", BlockID == boi, Replicate == 1) %>%
  ReshapeData(., data.type = "viability") %>%
  CalculateSynergy(method = "ZIP", correction = TRUE) %>%
  PlotSynergy(type = "all", save.file = FALSE)
```

```{r}
boi = c("TAS-102_AZD 5363")
moi = "ZIP"

tas_syn %>%
  mutate(boi = if_else(BlockID %in% boi, TRUE, FALSE)) %>% 
  filter(method == moi) %>%
  ggplot(aes(as.factor(boi), score_average, group=boi)) +
  geom_jitter(aes(color=line), width = 0.3) + 
  stat_summary(fun.y = 'mean', fun.ymin ='mean', fun.ymax='mean', colour = 'black', geom='crossbar', width=0.5) + 
  geom_signif(comparisons = list(c("TRUE", "FALSE")), method = "t.test") + 
  theme_classic() + 
  ggtitle(paste0(boi," method:", moi))

syn_mat %>% filter(line == "07", BlockID == boi, Replicate == 1) %>%
  ReshapeData(., data.type = "viability") %>%
  CalculateSynergy(method = "ZIP", correction = TRUE) %>%
  PlotSynergy(type = "all", save.file = FALSE)
```

```{r}
boi = c("TAS-102_Bosutinib")
moi = "ZIP"

tas_syn %>%
  mutate(boi = if_else(BlockID %in% boi, TRUE, FALSE)) %>% 
  filter(method == moi) %>%
  ggplot(aes(as.factor(boi), score_average, group=boi)) +
  geom_jitter(aes(color=line), width = 0.3) + 
  stat_summary(fun.y = 'mean', fun.ymin ='mean', fun.ymax='mean', colour = 'black', geom='crossbar', width=0.5) + 
  geom_signif(comparisons = list(c("TRUE", "FALSE")), method = "t.test") + 
  theme_classic() + 
  ggtitle(paste0(boi," method:", moi))

syn_mat %>% filter(line == "07", BlockID == boi, Replicate == 1) %>%
  ReshapeData(., data.type = "viability") %>%
  CalculateSynergy(method = "ZIP", correction = TRUE) %>%
  PlotSynergy(type = "all", save.file = FALSE)
```

#To-Dos
Add library data and look for enrichments. 

#Conclusion 

