---
title: "Check replicate quality TAS-102 Screen"
author: "Niklas Rindtorff"
output: 
   BiocStyle::html_document2:
    code_folding: hide
    df_print: paged
    toc_depth: 3
    toc_float: true
    fig_height: 6
    fig_width: 9.5
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=9.5, fig.height=6) 
# Set some global knitr options
library("knitr")
opts_chunk$set(tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=60), cache=TRUE, messages=FALSE)
```

#Load packages

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(pheatmap)
library(platetools)
library(viridis)
```

#Load data

```{r, warning=FALSE, message=FALSE}
##read ctg data files
ctg_data <- lapply(list.files('~/tas/data/replicate_1/', full.names=T, pattern = "TAS"), 
                   function(f) read_tsv(f, col_names=F, col_types = 'cci') %>%
                     `colnames<-`(c('screen', 'well', 'pcount'))) %>% bind_rows() %>%
  separate(screen, c("date", "operator", "mithras", "experiment_id"), sep = "_", remove = FALSE) %>%
  separate(experiment_id, c("experiment", "line", "TAS_concentration")) %>%
  #join compound annotation data
  full_join(., lapply(list.files('~/tas/data/annotations/', full.names=T), 
                   function(f) read_csv(f)) %>% bind_rows() %>%
  rename(well = destination.well) %>%
  select(drug, well, Source_Position, Source) %>%
  separate(well, c("row", "col"), sep = 1, remove = FALSE) %>%
  mutate(drug = if_else(is.na(drug), Source, drug),
         drug = if_else(well == "J19", "DMSO", drug),
         combination = if_else(col == "13", FALSE, TRUE)))

#read and combine error logs
# error_log <- lapply(list.files('~/tas/data/replicate_1/', full.names=T, pattern = "error"), 
#                    function(f) read_delim(f, "\t", escape_double = FALSE, trim_ws = TRUE) %>%
#                      `colnames<-`(c('screen', 'plate_error', 'well_error'))) %>% bind_rows()

#load annotation file 
# l11 <- lapply(list.files('~/tas/data/annotations/', full.names=T), 
#                    function(f) read_csv(f)) %>% bind_rows() %>%
#   
#   rename(well = destination.well) %>%
#   mutate(drug = if_else(is.na(drug), Source, drug)) %>%
#   select(drug, well, Source_Position) #concentration info is in Source.Position #todo
```
#Inspect screening data

First I want to gain a broad overview of the asssay plates. 

```{r, warning=FALSE}
tmp <- ctg_data

raw_grid(data = tmp$pcount,
           well = tmp$well,
           plate_id = tmp$screen,
           plate = 384,
          ncols = 6) + 
    theme_minimal() + 
    scale_fill_viridis() + 
    theme(aspect.ratio=0.66667)
```

#QC
Now I look at the distribution of controls in the first screening replicate.

```{r, warning=FALSE}
coi <- c("Stauro_500nM", "DMSO")

ctg_data %>%
  filter(drug %in% coi) %>%
  ggplot(aes(drug, pcount, color = combination)) +
  geom_jitter(alpha = 0.7) + 
  facet_grid(line ~ TAS_concentration) + 
  theme_bw()
```

The distribution seems to be quiet good. Only plate TAS-27-4 seems to have suffered a bit due to a delayed CTG readout. I now go on calculating the z-factors for each plate for combined and uncombined wells. 

```{r, warning=FALSE}
ctg_data %>%
  filter(drug %in% coi) %>%
  group_by(line, TAS_concentration, drug, combination) %>%
  summarise(sd = sd(pcount, na.rm = TRUE),
            mean = mean(pcount, na.rm = TRUE)) %>%
  group_by(line, TAS_concentration, combination) %>%
  summarise(zfactor = 1-((3*sum(sd))/abs(range(mean)[1]-range(mean)[2]))) %>%
  mutate(qc = if_else(zfactor < 0.5, FALSE, TRUE)) %>%
  ggplot(aes(combination, zfactor, color = qc)) +
  geom_point(alpha = 0.7, size = 3) + 
  facet_grid(line ~ TAS_concentration) + 
  theme_bw() + 
  geom_hline(yintercept = 0.5)
```

Also the z-factors are okay for all plates. Three plates have slight losses in assay quality, however, this seems still acceptable. I would continue 
